---
title: "STA 3100 Programming with Data: Assignment 010"
params:
  term: "Fall 2022"
  duedate: "2022-09-13"
  section: "4515 / 6736"
  classnumber: "25533 / 26666"
  name: "Camila Djurinsky Zapolski"
  ufid: "31303092"
  email: "djurinsky.camila@ufl.edu"
  total_points: 20
  grade: NA
output: 
  html_document: default
  pdf_document: default
---
TA comments:

No answer for question 1a on knitted file (note that TA does not read rmd file to grade). 1pt taken off.


1pt taken off for 3c for not using "probs" argument.


# A Random Angle

If the random variables $X$ and $Y$ are independent and uniformly
distributed on the unit interval (the interval from $0$ to $1$), then
the ordered pair $(X,Y)$ is uniformly distributed on the unit square
$A = \{(x,y) : 0 \le x \le 1, 0 \le y \le 1\}$.  Here is an example of
a single such point generated in R.  (Here I have set the seed for the
random number generator so that I get the same result each time this
code is run.)

```{r}
set.seed(324)
X <- runif(1)
Y <- runif(1)
```

With the seed set as above, the point generated is
$(X,Y) = (0.7290204, 0.4236485)$.
(If you change the seed or remove the `set.seed()` command altogether,
then you will get a different point.)

Now consider the random angle $\theta$ between the $x$-axis and the
segment joining the origin $(0,0)$ to the point $(X,Y)$.  Notice that
the tangent of this angle is $Y/X$ ("opposite over adjacent"; see
figure), so in our example the observed angle (in degrees) is
$$
\theta = \tan^{-1}(0.4236485/0.7290204) = 30.1617392.
$$



Suppose that we are interested in the distribution of the random angle
$\theta$. It is possible to derive the probability density of $\theta$
mathematically (see STA 4321), but even if we do not know how to do
this, we can still approximate the distribution by simulation.
To do this, we must draw many random observations from the
distribution of $\theta$, or equivalently, we must draw a large number
of independent $X$ and $Y$ values from the uniform distribution and
calculate the value of theta corresponding to each $(X,Y)$ pair (R's
vectorized functions make this easy). Then the probability of any
$\theta$-event can be approximated by calculating the proportion of
simulated $\theta$ values for which the event occurs.

# Exercises

1. (10 pts) Calculate 100,000 independent simulated observations of
   $\theta$ from the random experiment described in the introduction
   (do not print the values).
```{r setup}
X <- runif(100000)
Y <- runif(100000)
theta <- atan(Y/X)*(180/pi)

```
2. 
   a. (5 pts) With the $\theta$ values generated in exercise #1, use the
      ` histogram()` function (with `freq = FALSE`) to graph an
      estimate of the probability density of $\theta$.
   
```{r}

hist(theta, freq = FALSE)


```
   
   b. (5 pts) Based on the histogram, does the distribution appear to be
      symmetric or is it skewed?  Is it unimodal or multimodal or
      uniform (flat)?
   
    Based on the histogram, it would seem that the distribution is both symmetrical and unimodal.
   
   c. (5 pts) Based on the histogram, would you say that the angle is
      uniformly distributed from 0 to 90 degrees, i.e., does it appear
      that all values between 0 and 90 degrees are equally likely?
    
    No, the angle is normally distributed. This is shown by the fact that angles between 40 and 50 degrees are more likely than angles between 0 and 10 degrees or 70 to 80 degrees.
 
3. Using your simulated values, estimate:

   a. (5 pts) The mean of the distribution of $\theta$.
   
   
    The mean of the distribution is `r mean(theta)`
   
   b. (5 pts) The standard deviation of the distribution.
   
   
    The standard deviation of the distribution is `r sd(theta)`
   
   c. (5 pts) The lower quartile, median, and upper quartile of the distribution (use the `quantile()` function for this).

```{r}
thetaNums <- quantile(theta)
LQ <- thetaNums[2]
median <- thetaNums[3]
UQ <- thetaNums[4]

# should have been
# quantile(theta, probs = seq(0.25, 0.50, 0.75))
#quantile(theta, probs = seq(0.50, 0.75))
#quantile(theta, probs = seq(0.75, 1))
```   
    The lower quartile of the distribution is `r LQ`
    
    The median fo the distribution is `r median`
    
    The upper quartile of the distribution is `r UQ`

4. (10 pts) Use your simulated values to estimate the probability that $\theta$ is between 30 and 60 degrees.

```{r}
thetaSet <- (30 < theta & theta < 60)
prob <- mean(thetaSet) *100
```

    The probability that 30 < theta < 60 is `r prob` %
