---
title: "STA 3100 Programming with Data: Assignment 020"
params:
  term: "Fall 2022"
  duedate: "2022-09-29"
  section: "4515 / 6736"
  classnumber: "25533 / 26666"
  name: "Camila Djurinsky Zapolski"
  ufid: "31303092"
  email: "djurinsky.camila@ufl.edu"
  total_points: 20
  grade: NA
output: 
  html_document: default
  pdf_document: default
---

# Berkeley Admissions Data

The dataset `UCBAdmissions` is distributed with base R.  For the year
1973, applicants to the graduate programs of the six largest academic
departments at the University of California Berkeley are
cross-classified by admission status, gender, and the department to
which they applied.

```{r}
ftable(UCBAdmissions, row.vars = c("Admit", "Gender"), col.vars = "Dept")
```

Note: Some of the functions that I used in the class notes have
acquired new names.  The old names still work, but `margin.table()`
has been renamed `marginSums()` and `prop.table()` has been renamed
`proportions()`.  You can use either, but I would suggest adopting the
new function names.

## Exercises

1. (5 pts) Display the marginal table cross-classifying admission by
   gender (ignoring department).
   
```{r setup}
margin.table(UCBAdmissions, c("Admit", "Gender"))
```
2. (5 pts) Display the marginal table giving the proportions admitted
   and rejected by gender (ignoring department).  Express your results as
   percentages, rounded to one decimal place.  For whom is the
   admission rate higher overall, males or females?
   
```{r}
round(100 * prop.table(margin.table(UCBAdmissions, c("Admit", "Gender")), "Admit"), 1)
```
   Overall, the admission rate is higher for males.

3. (10 pts) Now display a table giving the proportion admitted by
   gender within each department. Express your results as percentages,
   rounded to one decimal place. How do these results compare to what
   you found in exercise 2?
```{r}
ftable(round(100 * prop.table(margin.table(UCBAdmissions, c("Admit", "Gender", "Dept")), "Dept"), 1), col.vars = c("Admit", "Gender"), row.vars = ("Dept"))
```
These results show the breakdown of admissions based on gender within each department. With this table a different conclusion can be drawn than from the table in exercise 2. In this table we see that the proportions of admission within each department differ, with departments A, B, and D departments having a greater admission of males, while departments C, E and F had a greater admission of females. Granted the difference in admissions also differed greatly, and although the total of overall male admissions was still greater than that of females, this table allows for the argument that females took the lead in the admissions of certain departments. 